Enforced Invariant Confluence

### 2018-03-31 doc/paper/sections/merge_reduction.tex
Minor tweaks.
Finished merge reduction section.

    \item
      For every $t \in T$, there exists some $s_t \in S$ such that for all $s
      \in s$, $t(s) = s \join s_t$. That is, every transaction $t$ is of the
      form $t(s) = s \join s_t$ for some constant $s_t$.

@@

% interactive is nice but requires user feedback; here we develop an alternate suff cond; to do so we introduce blah blah elimination
% merge elimination
% new sufficient condition
% example showing they're not equivalnet
% some discussion taht it's general

>>>>

    \item
      For every $t \in T$, there exists some $s_t \in S$ such that for all $s
      \in S$, $t(s) = s \join s_t$. That is, every transaction $t$ is of the
      form $t(s) = s \join s_t$ for some constant $s_t$.

@@

The proof of \thmref{LatticeProperty} involves proving that if criteria (1)
through (4) are met, then $O$ is \sTImergereducible{}. From this, it's
immediate that $O$ is \sTIconfluent{} by
\thmref{ReducibilityImpliesIconfluence}. To show that $O$ is
\sTImergereducible, we consider two merge-free reachable expressions
\[
  e_1 = t_1(\ldots(t_n(s_0))\ldots)
  \quad\text{and}\quad
  e_2 = u_1(\ldots(u_m(s_0))\ldots)
\]
Then, we inductively use criterion (3) with criteria (1) and (2) to show that
\[
  e_3 = t_1(\ldots(
          t_n(u_1(\ldots(u_m(s_0))\ldots))
        )\ldots)
\]
is reachable and evaluates to $e_1 \join e_2$.

\thmref{IclosureImpliesIconfluence} states that invariant-closure is a
sufficient condition for invariant-confluence, and \thmref{LatticeProperty}
states that criteria (1) -- (4) are sufficient conditions for
invariant-confluence. How do these sufficient conditions relate to one another?
Clearly, not all invariant-closed objects are semilattices, so
invariant-closure does not imply criteria (1) -- (4). Conversely, there are some
objects which satisfy criteria (1) -- (4) which are not invariant-closed. Here's
one example.

\begin{example}\examplelabel{TwoSets}
  Let $O = (\mathcal{P}(\nats), \cup)$ where $\mathcal{P}(\nats)$ is the power
  set of the natural numbers. Our start state $s_0 = \set{0}$ is the set of
  $0$. Let $t_Y(X) = X \cup Y$ be the transaction which unions $Y$ with its
  argument $X$. Our set $T = \setst{t_Y}{Y \subseteq \nats}$ of transactions
  consists of all possible $t_Y$. Our invariant $I$ consists of all non-empty
  sets $X$ that contain only even or only odd elements. That is, $I = \setst{X
  \subseteq 2\nats}{X \neq \emptyset} \cup \setst{X \subseteq 2\nats + 1}{X
  \neq \emptyset}$.

  Clearly, criteria (1), (2), and (4) are satisfied. Criterion (3) is also
  satisfied. If $s$ is an arbitrary state that satisfies $I$, then it is
  non-empty and contains, without loss of generality, only even integers. If
  $t_1$ and $t_2$ are arbitrary transactions such that $I(t_1(s))$ and
  $I(t_2(s))$, then $t_1(s)$ and $t_1(s)$ are also non-empty and contain only
  even integers. Thus, $t_1(s) \cup t_2(s)$ is clearly non-empty and contains
  only even integers, so $I(t_1(s) \join t_2(s))$.
  %
  However, $O$ is not \Iclosed{}. Let $s_1 = \set{0}$ and $s_2 = \set{1}$.
  Then, $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \cup s_2 = \set{0, 1}$,
  $\lnot I(s_3)$.
\end{example}

{\input{figures/merge_reducible}}

Invariant-closure is not necessary for invariant-confluence because it fails to
incorporate any notion of reachability. Criteria (1) -- (4) are also
unnecessary, but they can be used to prove that some non-invariant-closed
objects are invariant-confluent because the criteria \emph{do} incorporate
notions of reachability. In particular, criterion (3) is a slight variant of
invariant-closure; it also states that invariant-satisfying states should be
closed under merge. The fundamental difference is that criterion (3) restricts
its attention to the merge of two states which are \emph{reachable} from a
common ancestor state.

In \exampleref{TwoSets}, we saw this fundamental difference rear its head. $O$
is not $I$-closed because the union of an odd-only set with an even-only set is
a set with both odd and even integers. However, if we begin in an
invariant-satisfying state, we cannot reach both an odd-only and even-only set.
Criterion (3) is able to recognize this fact and conclude that $O$ is
invariant-confluent despite it not being invariant-closed.

% TODO: Discuss generality?
